Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-11 paper/comoving-followup.tex
intro

\begin{abstract}
% Context
Widely-separated ($\gtrsim 0.01$~pc), coeval stars are powerful tools for
testing star-formation model predictions, improving stellar atmosphere models,
and inferring the mass distribution of the Milky Way.
They likely outnumber the Galactic open clusters by a large margin, but are
still expected to share a common origin and chemistry.
The kinematic properties of comoving pairs---including bound wide binaries---
serve as potentiometers for the smooth tidal field and mass perturbation
spectrum of the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood using HIPPARCOS.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
false-positive rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own low-resolution radial-velocity survey of a
sample of the candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, relative to the kinematic and chemical
dispersions observed amongst typical field stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way.
The initial kinematic associations between stars from the same cluster are then
erased as the constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way (e.g., \citealt{todo}).
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children and preserved so that their initial association may be inferred
with precisely measured abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on not only on the precision of
the abundance measurements but also on the intrinsic chemical
dispersions of stars formed from the same birth clusters,
and the ability to tell that the chemical abundances of
two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
Recent works suggest that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
stellar masses and evolutionary stages.
Precise kinematic and color data from the \gaia\ mission
(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
existing membership catalogs and will lead to discoveries of many more open
clusters and young star associations (see, e.g., \citealt{Oh:2017}).
However, because they disrupt quickly (\smoh{a quantitative time scale?},
the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated ($\gtrsim 0.01~\pc$), coeval pairs or multiplets of stars.
These tracers include both bound and unbound wide-separation stellar systems; we
refer to these systems collectively as ``comoving stars'' because, at large
separations with present-day and near-future velocity uncertainties, it is
typically not possible to determine whether these systems are truly bound.
Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
dissolving or disrupted open clusters.
In any of these cases, it is likely that the stars formed from the same region
of a birth cloud, yet are separated enough that there has not been any
cross-pollution in the event of strong stellar winds or stellar evolution.

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
\citealt{Oh:2017,Oelkers:2017,Andrews:2017}).
Radial velocity follow-up is necessary to confirm pairs using these astrometric
catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$).
High-resolution spectroscopy is still needed to measure detailed chemical
abundances of these pairs, which may present interesting surprises; one
large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
high-resolution, high signal-to-noise spectra display large abundance
differences ($\approx 0.2~\dex$) that appear to have resulted from massive
accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
submitted).

In this work, we present a catalog of genuine comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars selected from
inferences using their astrometry alone (\citealt{Oh:2017}).
This work serves as a pilot study for data expected in upcoming data releases
from the \gaia\ mission.
We describe the data, observations, and spectroscopic reduction in
\sectionname~\ref{sec:data}.
If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
and \sectionname~\ref{sec:discussion} where we present and discuss our main
results.

>>>>

\begin{abstract}
% Context
Widely-separated ($\gtrsim 0.1$~pc), coeval stars are powerful tools for
testing star-formation model predictions, improving stellar atmosphere models,
and inferring the mass distribution of the Milky Way.
They likely outnumber the Galactic open clusters by a large margin, but are
still expected to share a common origin and chemistry.
The kinematic properties of comoving pairs---including bound wide binaries---
serve as potentiometers for the smooth tidal field and mass perturbation
spectrum of the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood using HIPPARCOS.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
false-positive rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own low-resolution radial-velocity survey of a
sample of the candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

Widely-separated, coeval pairs or multiplets of stars --- i.e. wide binaries,
multiplets, or former binaries or multiplets --- are important tracers of many
processes that formed and continue to shape the Milky Way.
For example, they have been used to place limits on the mass distribution and
population of massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to
study the dynamical outcomes of star formation processes (e.g.,
\citealt{Reipurth:2012}), and to calibrate stellar models and inferences across
the main sequence (e.g., \citealt{Brewer:2016}).
These examples represent a small subset of a much broader range of applications
(see, e.g., \citealt{Chaname:2007}).

At smaller separations ($\lesssim 0.1~\pc$), comoving pairs of stars are likely
bound wide binaries and are therefore coeval.
At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
tenuously bound stars, recently disrupted binaries, moving groups or dynamical
features, and the final products of dissolving stellar associations or open
clusters.
However, even at separations of $\sim 0.1~\pc$, we typically lack the precision
to determine the binding energy of a system and therefore can't determine its
true dynamical state; because of this ambiguity, we refer to both bound and
unbound systems collectively as ``comoving stars.''

Comoving stars are identified by searching for approximately cospatial stars
that appear to be comoving either in full-space or in sky-projected velocities.
Primarily thanks to large astrometric surveys and catalogs
(\citealt{ESA:1997,Lepine:2005,Gaia-Collaboration:2016}) and radial-velocity
surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving pairs
(\citealt{Shaya:2011}) and over 10,000 candidate systems
(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,Oelkers:2017,
Andrews:2017}) have already been discovered.
Radial velocity follow-up is still needed to confirm the majority of these
pairs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).

It has long been recognized that the observed period or separation distribution
of binary stars encapsulates information about star formation and dynamical
processes of the Galaxy (\citealt{Opik:1924}).
[If the initial distribution power law, break and steepens from perturbations]
[a bit of theory about wide binaries...predictions for wide binary separations?]
(Weinberg87)

Using candidate comoving pairs selected based on their proper motions alone,
several groups have measured the separation distribution of comoving stars at
angular separations $\Delta \theta \lesssim 1^\circ$ where the false-positive
rate is expected to be lower (\citealt{Chaname:2004,Lepine:2007,Sesar:2008});
these are likely predominantly wide binaries.
In most cases, the angular or projected separation distribution is fit using a
decreasing power-law, and the power-law index, $\alpha$, is found to be $\alpha
\approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
0.65~\pc$ at a distance of $150~\pc$).
Tentative reports of a steepening or cutoff of the separation distribution at
larger separations have been claimed (e.g., \citealt{Yoo:2004,Quinn:2009}), but
these claims are limited by small numbers of confirmed pairs at large
separations.

While the separation distribution of wide binaries appears to and is expected to
fall off rapidly with increasing separation, predictions for the separation
distribution of \emph{comoving stars} are less clear.
Early work suggested ionized pairs don't matter much...
[But including tidal field...] (jiang and tremaine)
[But still haven't included perturbers...]
[Tentative evidence of this population in (shaya and olling)]
[Need larger samples of confirmed large-separation pairs]

In this work, we present a catalog of confirmed comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars selected from
inferences using their astrometry alone (\citealt{Oh:2017}).
This work serves as a pilot study for data expected in upcoming data releases
from the \gaia\ mission.
We describe the data, observations, and spectroscopic reduction in
\sectionname~\ref{sec:data}.
If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
and \sectionname~\ref{sec:discussion} where we present and discuss our main
results.

% The majority of stars are thought to form in dynamically cold, chemically
% homogeneous clusters, relative to the kinematic and chemical
% dispersions observed amongst typical field stars in the Galactic disk.
% Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
% progenitor gas cloud disperses quickly (cite), it is typically assumed that
% stars formed in the same birth cluster have the same initial chemistry
% (e.g., cite + \citealt{Feng:2014}) and same ages.
% These birth clusters are typically born un- or loosely-bound and therefore
% quickly succumb to the tidal field of the Milky Way.
% The initial kinematic associations between stars from the same cluster are then
% erased as the constituent stars mix into the general population of disk stars.
% However, the stars do preserve their initial associations in age, chemical
% abundance patterns, and dynamical invariants (e.g., orbital actions).
% It has therefore been long presumed that ``chemical tagging'' and chemodynamical
% models of the Galactic disk will allow for inferences about the formation and
% evolution of the Milky Way (e.g., \citealt{todo}).
% That is, the chemical environment of a birth cluster is likely imprinted on its
% stellar children and preserved so that their initial association may be inferred
% with precisely measured abundance patterns (of $\gtrsim 10$ species).

% The feasibility of chemical tagging depends on not only on the precision of
% the abundance measurements but also on the intrinsic chemical
% dispersions of stars formed from the same birth clusters,
% and the ability to tell that the chemical abundances of
% two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
% Recent works suggest that, at least for red giant stars, the abundance
% distributions of 20 elements in a handful of open clusters included in the
% \acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
% with having intrinsic dispersions smaller than the abundance measurement
% precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
% intrinsic dispersions are not known.
% This demonstration of chemical uniformity in open clusters is promising for
% chemical tagging.
% However, it was also recently shown that $\approx 0.3\%$ of random pairings of
% Galactic field red giant stars have abundance patterns that are
% indistinguishable at the same level (``doppelgangers''), even with 20 chemical
% species (\citealt{Ness:2017}).

% Are these properties of open clusters universal, or limited to the few clusters
% presently studied?
% Do the chemical properties of open clusters depend on the star formation rate or
% efficiency, which may evolve with time?
% Ideal tests of the chemical homogeneity assumptions would use many more open
% clusters with varied cluster masses and member stars that span a large range in
% stellar masses and evolutionary stages.
% Precise kinematic and color data from the \gaia\ mission
% (\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
% existing membership catalogs and will lead to discoveries of many more open
% clusters and young star associations (see, e.g., \citealt{Oh:2017}).
% However, because they disrupt quickly (\smoh{a quantitative time scale?},
% the present-day open cluster population
% is biased towards younger or more massive clusters.

% High-resolution spectroscopy is still needed to measure detailed chemical
% abundances of these pairs, which may present interesting surprises; one
% large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
% discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
% high-resolution, high signal-to-noise spectra display large abundance
% differences ($\approx 0.2~\dex$) that appear to have resulted from massive
% accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
% submitted).
